---
title: "Incarceration Analysis"
author: "Sarah Mun"
date: "2/25/2022"
output: html_document
---

```{r setup, include=FALSE}
library(dplyr)
library(tidyverse)
library(ggplot2)
library(plotly)
library(leaflet)

source("../source/analysis.R")
source("../source/line_plot.R")
source("../source/scatterplot_matrix.R")
source("../source/map.R")
```

## Introduction

This report summarizes my exploration of the Vera Institute's data set, which analyzes the population of incarcerated peoples between the years of 1970 and 2016. This analysis also allows us to take a closer look at how different racial demographics make up the U.S. incarceration population. 

## Summary Information

In my exploration of the data set, I wanted to get an understanding of how the various demographics make up the incarcerated population specifically in the year 2016 - since this is the most recent year in the data set that had the necessary values. I found that the location with the highest incarceration population in 2016 was `r summary_info$location_highest_2016` compared to `r summary_info$location_highest_1970` in 1970. The average total incarcerated population in 2016 was `r summary_info$avg_total`. When we look at these demographics my race, we get that the AAPI population was `r summary_info$avg_aapi`, the Black population was `r summary_info$avg_black`, the Latinx population was `r summary_info$avg_latix`, the Native population was `r summary_info$avg_native`, the White population was `r summary_info$avg_white`, and the "other races" population was `r summary_info$avg_other`. Of all the racial demographics listed, it appears that the White demographic makes up the highest amount of the population of incarcerated people on average. Not only does this data set consider racial demographics, but it also takes note of sex. The ratio between the average female and male population in prisons for 2016 was `r summary_info$ratio_male_female`.

```{r, echo=FALSE}
summary_info
```

## Chart 1: Comparing Black and White Demographics in 2016.

This chart shows how the incarcerated population of these two demographics have changed over time in King County. I chose these two demographics specifically because people of color, especially those in Black communities, are disproportionately affected by racial profiling and are five time more likely to be imprisoned than those in White communities.

```{r, echo=FALSE}
black_vs_white
```

## Chart 2: Incarceration Rates Between 1970 and 2016 in King County

This chart shows how the total prison population in King County has changed over the course of the data set. According to the Vera Institute, this value is defined as the average daily number of people held in prison at any time during the year. 
```{r, echo=FALSE}
total
```

## Chart 3: Incarceration Across the U.S. in 2016

This chart shows how the prison population varies across the U.S.. From this graph, we can see that states with large populations and large cities have higher incarcerated populations than more rural areas. 
```{r, echo=FALSE}
map
```
