---
title: "Incarceration Analysis"
author: "Sarah Mun"
date: "2/25/2022"
output: html_document
---

```{r setup, include=FALSE}
library(dplyr)
library(tidyverse)
library(ggplot2)
library(plotly)
library(leaflet)

setwd("~/Desktop/INFO-201-coding/a3-sarah02mun/docs")

source("../source/line_plot.R")
source("../source/scatterplot_matrix.R")
source("../source/map.R")
```

## Introduction

This report summarizes my exploration of the Vera Institute's data set, which analyzes the population of incarcerated peoples between the years of 1970 and 2018. Through this analysis, we can take a closer look at how these 

```{r cars}
summary(cars)
```

## Chart 1: Comparing Black and White Demographics in 2018.

This chart shows how the incarcerated population of these two demographics have changed over time in King County. I chose these two demographics specifically because people of color, especially those in Black communities, are disproportionately affected by racial profiling and are five time more likely to be imprisoned than those in White communities. This graph shows how this trend is manifested in the King County prison system.

```{r, echo=FALSE}
black_vs_white
```

## Chart 2: Incarceration Rates Between 1970 and 2018 in King County

This chart shows how the total jail population in King County has changed over the course of the data set. According to the Vera Institute, this value is defined as the average daily number of people held in jail at any time during the year. 
```{r, echo=FALSE}
matrix
```

## Chart 3: Incarceration Across the U.S. in 2018
```{r, echo=FALSE}
incarceration_2018_map
```
